### 2020-07-17 paper.tex
tweaks to sec 3 and 4

In this section, we introduce the Gen language~\citep{gen-pldi} for probabilistic programs, and present a technique for automating the implementation of involutive MCMC algorithms when the model $p$ and auxiliary distribution $q_x$ are both represented as
Gen programs.

\subsubsection{A Language for Probabilistic Programs with Traces}

The Gen probabilistic programming language augments the syntax of Julia~\citep{julia}
with a single new construct, the \texttt{\{a\} $\sim$ d} expression, for making a \textit{named} random choice. An execution trace of a Gen probabilistic program can be sampled by running the program according to Julia's usual semantics, and upon encountering an expression of the form \texttt{\{a\} $\sim$ d}, (i) evaluating the \textit{address expression} \texttt{a} to obtain an address $a \in \mathcal{A}$; (ii) evaluating the \textit{distribution expression} $d$ to obtain a distribution $d$ over the measurable space $V_a$; (iii) sampling a value $x_a \sim d$ and adding the mapping $\{a \mapsto x_a\}$ to a dictionary called the program trace; and (iv) returning the sampled value $x_a$ to the program, to continue execution. When execution terminates, the program trace has accumulated a mapping for each random choice encountered during the program's execution.

As a concrete example, consider Figure~\ref{fig:mixture}a, which defines a Gen program for a Gaussian mixture model with an unknown number of components. The program \texttt{p} accepts as input an integer \texttt{n}, a number of data points. Each $\texttt{n}$ defines
a distinct distribution over dictionaries.

>>>>

In this section, we introduce a probabilistic programming language from the Gen probabilistic programming system~\citep{cusumano2019gen}, and present a technique for automating the implementation of involutive MCMC algorithms when the model $p$ and auxiliary distribution $q_x$ are both represented as
probabilistic programs in this language.
But the technique is not limited to the Gen system---we also provide the implementation of a minimal probabilistic programming language in PyTorch that supports the automation technique.

\subsubsection{A Probabilistic Programming Language}

Our probabilistic programming language augments the syntax of Julia~\citep{bezanson2017julia}
with a new construct, the \texttt{\{a\} $\sim$ d} expression, for making a \textit{named} random choice. An execution trace of a Gen probabilistic program can be sampled by running the program according to Julia's usual semantics, and upon encountering an expression of the form \texttt{\{a\} $\sim$ d}, (i) evaluating the \textit{address expression} \texttt{a} to obtain an address $a \in \mathcal{A}$; (ii) evaluating the \textit{distribution expression} $d$ to obtain a distribution $d$ over the measurable space $V_a$; (iii) sampling a value $x_a \sim d$ and adding the mapping $\{a \mapsto x_a\}$ to a dictionary called the program trace; and (iv) returning the sampled value $x_a$ to the program, to continue execution. When execution terminates, the program trace has accumulated a mapping for each random choice encountered during the program's execution.

As a concrete example, consider the probabilistic program in Figure~\ref{fig:mixture}a, which defines a Gaussian mixture model with an unknown number of components:
\noindent
\begin{center}
\begin{tabular}{c}
{\begin{lstlisting}[basicstyle=\small\ttfamily]
@gen function p(n::Int)
  k ~ poisson_plus_one(1)
  means = [
    ({(:mu, j)} ~ normal(0, 10)) for j in 1:k]
  vars = [
    ({(:var, j)} ~ inv_gamma(1, 10)) for j in 1:k]
  weights ~ dirichlet([2.0 for j in 1:k])
  for i in 1:n
    {(:x, i)} ~ mixture_of_normals(
        weights, means, vars)
  end
end
\end{lstlisting}}
\end{tabular}
\end{center}
The program \texttt{p} accepts as input an integer \texttt{n}, a number of data points. Each $\texttt{n}$ defines
a distinct distribution over dictionaries.
